ROV report

### 2020-03-23 analysis/reports/rov.Rmd
writing and dnew lots

all_di <- filter(all_di, s0 > 2)
```

@@

```{r size of fs}
ggplot(filter(all_di, singletons == F), aes(x = log(s0), y = log(n0), color = log_nparts)) +
  geom_point(alpha = .5) +
  theme_bw() +
  theme(legend.position = "top") + scale_color_viridis_c(option = "magma", end = .9, begin = .1)

@@

bin_rov <- function(rov_vector, nbins, quantiles = F) {

rov_long <- rov_long %>%
  group_by(rov_index) %>%
  mutate(binned_val = bin_rov(val, nbins =5 ),
         qbinned_val = bin_rov(val, nbins = 5, quantiles = T)) %>%
  ungroup()

@@

for(i in 1:length(unique(rov_long$rov_index))) {
  rov_nparts_plots[[i]] <- ggplot(filter(rov_long, rov_index == unique(rov_long$rov_index)[i]), aes(log_nparts, val, color = dat)) +
    geom_point(alpha = .2) +
    theme_bw() +
    ggtitle(unique(rov_long$rov_index)[i]) +
    geom_vline(xintercept = 10) +
    scale_color_viridis_d() +
    theme(legend.position = "none")
}

@@

If we bin nparts:

@@

for(i in 1:length(unique(rov_long$rov_index))) {
  rov_nparts_binned_plots[[i]] <- ggplot(filter(rov_long, rov_index == unique(rov_long$rov_index)[i]), aes(log_nparts, val, color = as.factor(thresholded_nparts))) +
    geom_point(alpha = .1) +
    theme_bw() +
    ggtitle(unique(rov_long$rov_index)[i]) +
    scale_color_viridis_d() +
    theme(legend.position = "none")
}

for(i in 1:length(rov_nparts_binned_plots)) {
  print(rov_nparts_binned_plots[[i]]) 
}
```

@@

@@

>>>>

all_di <- filter(all_di, s0 > 2, dat != "portal_plants", !singletons)

bin_rov <- function(rov_vector, nbins, quantiles = F, as.fact = F) {

  if(as.fact) {
    binned_vect <- factor(binned_vect, ordered = T, labels = round(cutoffs[2:length(cutoffs)],digits = 1))
  }

rov_long <- rov_long %>%
  group_by(rov_index) %>%
  mutate(binned_val = bin_rov(val, nbins =5 ),
         qbinned_val = bin_rov(val, nbins = 5, quantiles = T),
         qbinned_lognparts = bin_rov(log_nparts, nbins = 10, quantiles = T, as.fact = T)) %>%
  ungroup()

all_di <- all_di %>%
  mutate(qbinned_lognparts = bin_rov(log_nparts, nbins = 10, quantiles = T, as.fact = T))

@@

```{r size of fs}
ggplot(filter(all_di, singletons == F), aes(x = log(s0), y = log(n0), color = log(log_nparts))) +
  geom_point(alpha = .5) +
  theme_bw() +
  theme(legend.position = "top") + scale_color_viridis_c()

fs_size_panels <- list(
  ggplot(filter(all_di, singletons == F), aes(x = log(s0), y = log_nparts)) +
    theme_bw() +
    geom_point(),
  ggplot(filter(all_di, singletons == F), aes(x = log(n0), y = log_nparts)) +
    theme_bw() +
    geom_point(),
  ggplot(filter(all_di, singletons == F), aes(x = log(n0 / s0), y = log_nparts)) +
    theme_bw() +
    geom_point()
)

gridExtra::grid.arrange(grobs = fs_size_panels)

@@

@@

for(i in 1:length(unique(rov_long$rov_index))) {
  rov_nparts_plots[[i]] <- ggplot(filter(rov_long, rov_index == unique(rov_long$rov_index)[i]), aes(log_nparts, val, color = dat)) +
    geom_point(alpha = .2) +
    theme_bw() +
    ggtitle(unique(rov_long$rov_index)[i]) +
    scale_color_viridis_d() +
    theme(legend.position = "none")
}

@@

@@

for(i in 1:length(unique(rov_long$rov_index))) {
  rov_nparts_binned_plots[[i]] <- ggplot(filter(rov_long, rov_index == unique(rov_long$rov_index)[i]), aes(log_nparts, val, color = log(n0/s0))) +
    geom_point(alpha = .3) +
    theme_bw() +
    ggtitle(unique(rov_long$rov_index)[i]) +
    scale_color_viridis_c(end = .9) +
    theme(legend.position = "left")
}

for(i in 1:length(rov_nparts_binned_plots)) {
  print(rov_nparts_binned_plots[[i]]) 
}

ggplot(rov_long, aes(log_nparts, val, color = log(n0/s0))) +
  geom_point(alpha = .3) +
  theme_bw() +
  scale_color_viridis_c() +
  theme(legend.position = "left") +
  facet_wrap(vars(rov_index), scales = "free_y")
```

```{r qbinned nparts ROV}

rov_nparts_qbinned_plots <- list()

for(i in 1:length(unique(rov_long$rov_index))) {
  rov_nparts_qbinned_plots[[i]] <- ggplot(filter(rov_long, rov_index == unique(rov_long$rov_index)[i]), aes(qbinned_lognparts, val, color = log(n0 / s0))) +
    geom_jitter(alpha = .1) +
    theme_bw() +
    ggtitle(unique(rov_long$rov_index)[i]) +
    scale_color_viridis_c() +
    theme(legend.position = "none")
}

for(i in 1:length(rov_nparts_qbinned_plots)) {
  print(rov_nparts_qbinned_plots[[i]]) 
}
```

Remove low N/S:

```{r qbinned nparts ROV}

rov_nparts_qbinned_plots_highn <- list()

for(i in 1:length(unique(rov_long$rov_index))) {
  rov_nparts_qbinned_plots_highn[[i]] <- ggplot(filter(rov_long, rov_index == unique(rov_long$rov_index)[i], log(n0 / s0) > .7), aes(qbinned_lognparts, val, color = log(n0/s0))) +
    geom_jitter(alpha = .1) +
    theme_bw() +
    ggtitle(unique(rov_long$rov_index)[i]) +
    scale_color_viridis_c() +
    theme(legend.position = "none")
}

for(i in 1:length(rov_nparts_qbinned_plots_highn)) {
  print(rov_nparts_qbinned_plots_highn[[i]]) 
}
```

@@

Binned by quantile:

```{r qbinned percentiles}

ggplot(filter(all_di, !singletons, s0 > 2), aes(x = skew_percentile)) +
  geom_histogram(bins = 100, boundary = 0) +
  theme_bw() +
  facet_wrap(vars(qbinned_lognparts), scales = "free_y", nrow = 2)

ggplot(filter(all_di, !singletons), aes(x = simpson_percentile)) +
  geom_histogram(bins = 100, boundary = 0) +
  theme_bw() +
  facet_wrap(vars(qbinned_lognparts), scales = "free_y", nrow = 2)

@@

```{r report stats}

# proportion of %iles above 95/below 5
# skew
mean(all_di$skew_percentile >= 95)
# simpson
mean(all_di$simpson_percentile <= 5)

# by nparts

nparts_summary <- all_di %>%
  group_by(qbinned_lognparts) %>%
  summarize(high_skew = mean(skew_percentile >= 95),
            low_simpson = mean(simpson_percentile <= 5)) %>%
  ungroup()

nparts_summary

eighty_smallest <- filter(all_di, log_nparts <= 8.6)
mean(eighty_smallest$skew_percentile >= 95)

twenty_largest <- filter(all_di, log_nparts >= 8.6)
mean(twenty_largest$skew_percentile >= 95)

fifty_smallest <- filter(all_di, log_nparts <= median(log_nparts))
mean(fifty_smallest$simpson_percentile <= 5)
fifty_largest <- filter(all_di, log_nparts >= median(log_nparts))
mean(fifty_largest$simpson_percentile <= 5)
